---
title: "Models for tuition increases at Chicago Free School"
output:
  pdf_document: default
  html_notebook: default
---


```{r,echo=FALSE}
knitr::opts_chunk$set(echo=FALSE,results='asis')
library(readxl)
library(scales)
tuitions <- read_excel("tuitions.xlsx", 
    sheet = "Sheet 1", skip = 3)
tuitions$`family income`[64:70]<-as.character(seq(130000,to=160000,by=5000))
```

```{r,echo=FALSE,results='hide'}
#incomes<-1:130
calc_tuitions <- function(max_income=110,min_income=20,max_tuition=13.5, min_tuition=1.650,squared_term=0,percents_slope=NA,combine=FALSE,brackets=FALSE,bracket_boundaries=NA,bracket_percents=NA,
                          rounding=0) {

  

max_percent = max_tuition/max_income*100
min_percent = min_tuition/min_income*100

if(is.na(percents_slope)) {percents_slope = (max_percent-min_percent)/(max_income-min_income)}

incomes<-as.numeric(tuitions$`family income`)/1000
incomes[is.na(incomes)]=130
if(brackets){
  percents=vector(mode="numeric",length=length(incomes))
  for (i in seq(from=length(bracket_boundaries),to=1,by=-1)){
  #  print(bracket_boundaries[i])
  #  print(incomes<bracket_boundaries[i])
  percents[incomes<bracket_boundaries[i]]<-bracket_percents[i] 
  }
  percents[incomes>=bracket_boundaries[length(bracket_boundaries)]]<-bracket_percents[length(bracket_percents)]
}   else {
percents<-(incomes-min_income)*percents_slope + (incomes-min_income)*(incomes-min_income)*squared_term + min_percent
}




tuition<-pmin(max_tuition,pmax(min_tuition,incomes*percents/100))

if(rounding){
  tuition=round(tuition/rounding)*rounding
}

tuition[!is.na(tuitions$X__1=="sibling")]=tuition[!is.na(tuitions$X__1=="sibling")]/2


monthly_increase=(tuition-as.numeric(tuitions$tuition)/1000)*100

outputs<-list(tuition=tuition,monthly_increase=monthly_increase,
              percents=percents,
              incomes=incomes)
return(outputs)
}
```


```{r}
make_plots <- function(name,tuition,monthly_increase,incomes,tuitions){
#list2env(outputs,envir=environment())

  
par(mfrow=c(2,2))

plot(incomes,tuition,col='red',pch=18,xlab='Family income (thousands)',ylab='Annual tuition (thousands)',main=sprintf("Current tuition \n vs %s", name))
points(incomes,as.numeric(tuitions$tuition)/1000,col='black',pch=18)
plot(incomes,monthly_increase,pch=18,xlab='Family income (thousands)',ylab='Monthly tuition increase ($)',main="Tuition increases",ylim=c(0,110))

plot(incomes,monthly_increase/incomes,pch=18,xlab='Family income (thousands)',ylab='Increase as % of income',main=paste("Tuition increases \n as percent of income"),ylim=c(0,2))

plot(incomes,monthly_increase/as.numeric(tuitions$tuition)*1000,pch=18,xlab='Family income (thousands)',ylab='Increase as % ',main=paste("Tuition increases as \n percent of current tuition"),ylim=c(0,25))

sprintf('The mean tuition under this proposal is %d.',  round(mean(outputs$tuition)*1000))  

}
```


```{r}
outputs<-calc_tuitions()
```

## Current tuition vs income

```{r,echo=FALSE,results='hide'}
outputs<-calc_tuitions(min_tuition = 1.50,max_tuition=12,brackets=TRUE,bracket_boundaries=c(20.,40.,60.,80,100),bracket_percents = c(0,8.,9,10,11,12)+0,rounding=.25)  # Exactly reproduces current tuition

plot(outputs$incomes,as.numeric(tuitions$tuition)/1000,col='black',pch=18,xlab='Family income (thousands)',ylab='Annual tuition (thousands)',main="Current tuition")

```


The current mean tuition is `r round(mean(as.numeric(tuitions$tuition)))`.


\newpage
# Model 1: Carissa's Proposal

```
income	tuition assessment
under $20k	min = $1650
$20-40k	          9%
$40-60k	          10%
$60-80k	          11%
$80-100k	        12%
$100k+	          13%



```

Note that in this proposal, families earning \$1 less than the \$20,000 minimum will pay the minimum tuition of \$1,650. However, families earning \$1 above \$20,000 will pay at the 9\% rate, which is \$1,800. Currently these families pay at the minimum tuition of $1,500. We may wish to make an adjustment to fix this gap.

```{r,echo=FALSE}
par(mfrow=c(2,2))
outputs<-calc_tuitions(min_tuition = 1.650,max_tuition=13.5,brackets=TRUE,bracket_boundaries=c(20.,40.,60.,80,100),bracket_percents = c(0,8.,9,10,11,12)+1,rounding=.1)
make_plots("Carissa's proposal",outputs$tuition,outputs$monthly_increase,outputs$incomes,tuitions)
#make_plots("Carissa's proposal",outputs$floor_tuition,outputs$monthly_increase_floor,outputs$incomes,tuitions)


```



\newpage

# Model 2: 10\% increase across the board

This proposal has the same gap at the minimum income as Model 1 above.


```
income	tuition assessment
under $20k	min = $1650
$20-40k	          8.8%
$40-60k	          9.9%
$60-80k	          11%
$80-100k	        12.1%
$100k+	          13.2%



```


```{r,echo=FALSE}
par(mfrow=c(2,2))
outputs<-calc_tuitions(min_tuition = 1.650, brackets=TRUE,bracket_boundaries=c(20.,40.,60.,80,100),bracket_percents = c(0,8.,9,10,11,12)*1.1,rounding=.1)
make_plots("10% proposal",outputs$tuition,outputs$monthly_increase,outputs$incomes,tuitions)
#make_plots("Carissa's proposal",outputs$floor_tuition,outputs$monthly_increase_floor,outputs$incomes,tuitions)


```

\newpage




# Model 3: No-brackets proposal

In this proposal, tuition within the bracket range is calculated according to this formula: $$\text{tuition}=\frac{\text{income}-\$20,000}{\$84,000}\times 4.75\% + 8.25\%$$.

Corresponding brackets:

```
income	tuition assessment
under $20k	min = $1650
$20-40k	    8.25%-9.38%
$40-60k	    9.38%-10.51%
$60-80k	    10.51%-11.64%
$80-100k	  11.64%-12.77%
$100k-104k	12.77%-13%
```

```{r}
outputs<-calc_tuitions(max_income = 104,max_tuition=13.5,min_tuition=1.650,combine=FALSE,rounding=FALSE)

make_plots("No-bracket proposal",outputs$tuition,outputs$monthly_increase,outputs$incomes,tuitions)

```


\newpage




# Model 4: Combined proposal

In this proposal, tuition is the average of the tuitions calculated with the 10% across the board proposal and the no-brackets proposal. The idea is to use this proposal for one year, and then the following year go to the no-brackets proposal.


```{r}
outputs1<-calc_tuitions(max_income = 104,max_tuition=13.5,combine=FALSE,rounding=FALSE)
outputs2<-calc_tuitions(min_tuition = 1.650, brackets=TRUE,bracket_boundaries=c(20.,40.,60.,80,100),bracket_percents = c(0,8.,9,10,11,12)*1.1,rounding=FALSE)
outputs1$tuition<-(outputs1$tuition+outputs2$tuition)/2
outputs1$tuition<-ceiling(outputs1$tuition/.1)*.1 # do rounding
outputs1$monthly_increase<-(outputs$tuition-as.numeric(tuitions$tuition)/1000)*100

outputs1$monthly_increase<-(outputs1$monthly_increase+outputs2$monthly_increase)/2

make_plots("combined proposal",outputs1$tuition,outputs1$monthly_increase,outputs1$incomes,tuitions)

```



```


